\newpage
\chapter{Introduction}
\label{Intro}

The Standard Model (SM), developed in the early 1970s, encapsulates our best understanding of the fundamental particles and their interactions. It has demonstrated huge successes in providing experimental predictions which have been confirmed by the experimental observations over time. The discovery of the Higgs boson by the ATLAS and CMS experiments in 2012 at the Large Hadron Collider (LHC) was a breakthrough for the experimental tests of the SM. Currently, precise measurements of the Higgs boson couplings to fermions and bosons, mass and cross-section are performed with the Run-1 and Run-2 dataset collected with both ATLAS and CMS detectors, while the triple Higgs self-coupling, $\lambda_{HHH}$, still resist to physicists. The $\lambda_{HHH}$ is present in the SM and its value correspond to $m_{H}^2/2v$. This parameter controls the shape of the Higgs potential. The only direct way to measure this coupling is through the Higgs boson pair production process (HH). This process can be produced at the LHC mainly via gluon-gluon fusion (ggF) through destructive interference of two Feynman diagrams involving to-quark loops and the triple Higgs self-interaction, thus the possibility to access to it and probe the structure of the Higgs potential. At centre-of-mass energy of 13 TeV, the cross-section of the Higgs boson pair production is $33.53_{-6.0\%}^{+4.3\%}$ fb. The rate of Higgs boson pair production is 1000 smaller than the single Higgs production. Because of its small cross-section, the observation of this process and the measurement of the Higgs self-coupling is challenging and one of the main goals for the High Luminosity LHC (HL-LHC). 
Since Higgs is not a stable particle, the di-Higgs event is reconstructed through Higgs bosons decay products. The HH decay channels are the combination of single Higgs decay channels. At the Run-2 of LHC, the most sensitive decay channels are HH$\to$\bbbb, HH$\to$\bbtt, HH$\to$\bbyy and HH$\to\bar{b}bVV$. 
\\

If SM expectation hold, the observation of Higgs boson pair production is not possible with the currently available data, while many Beyond Standard Model (BSM) theories predict enhancements to its cross-section either through new resonances (resonant production) or non-resonant production. The non-resonant production assumes new physics anomalies that predict a different value for the Higgs self-coupling, this deviation from the SM predicted value is quantified with $\kappa_{\lambda} = \frac{\lambda^{new physics}}{\lambda^{SM}}$. Searches for non-resonant and resonant di-Higgs production were performed by the ATLAS and CMS collaborations in several decay channels using about 36 \ifb of pp collisions data from 2015-2016 at a centre-of-mass energy of 13 TeV. The ATLAS statistical combination of the \bbbb, \bbtt and \bbyy channels set an observed (expected) upper limit on the non-resonant HH production cross-section of around 6.9 (10) times the SM expectation and a constraint of the \kl parameters of [-5, 12] ([-5.8, 12]) at 95\% C.L. The CMS statistical combination of the same channels set an observed (expected) upper limit on the non-resonant HH production cross-section of around 22.2 (12.8) times the SM expectation and a constraint on \kl of [-11.8, 18.8] ([-7.1, 13.6]) at 95\% C.L. These results show that the sensitivity to this process is limited with statistics used, making it a flagship analysis for the HL-LHC. It is nevertheless important to continue exploring HH production with the increase of the luminosity in data 2015-2018 (139 \ifb) and to improve analysis techniques for this process. \\

This thesis focus on the search for the non-resonant Higgs boson pair production in the \bbyy final state, exploring the decay where one Higgs decays to two $b$-quark and the other to two photons, and the measurement of the self-coupling quantifier \kl using the full Run-2 data collected by ATLAS detector. The branching of this decay channel is around 0.3\%, while even with the low branching ratio it still a "golden" channel since it compromises between the H$\to\bar{b}b$ large branching ratio and the well-measured photon energy which makes it one of the most sensitive channels. This channel was already explored in early Run-2 using 36 \ifb of pp collisions. The full Run-2 analysis aims to improve the 36 \ifb results on top of the luminosity increase by implementing new analysis techniques.

This thesis is structured as follows: Chapter \ref{chap1} presents an overview of the SM and its particles, the Brout-Englert-Higgs mechanism, Higgs production, its decay at LHC and summary of the Higgs discovery, as well as a summary of the Higgs pair production and decay channels and the 36\ifb results. A description of the LHC and ATLAS detector, as well as the methods used to reconstruct the physics object, are presented in Chapter \ref{LHC&ATLAS}. To reconstruct the \HHyybb event an excellent knowledge of the photon, and therefore of the detector, is crucial. Chapter \ref{gamma} focus on the different methods used to reconstruct photon and a new photon identification algorithm is presented in this chapter. Similarly to photons, jets and $b$-jets are mandatory for \HHyybb analysis. Jets reconstruction and tagging as well as a specific calibration of the $b$-jet energy is presented in Chapter \ref{Jet}. The complete description of the \HHyybb analysis, including signal and background Monte Carlo simulations, objects and events selection, event categorization, background modelling, systematic uncertainties, statistical interpretation and results as well as comparison to the \HHyybb 36 \ifb results, \bbbb, \bbtt and the 139 \ifb CMS \HHyybb results, is given in Chapter \ref{HHyybb}. Finally, a description of the prospective for Run-3 analysis and the HL-LHC is given in Chapter \ref{HL-LHC}. 
